---
title: "NYC Crime Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

### Accessing the data. 

The data was accessed through the SODA Consumer API on 17th November, 2018. Click [here](https://dev.socrata.com/consumers/getting-started.html) to read more about the API. 

The `tidyverse`, `Rsocrata` and `lubridate` libraries are required to connect with the API and subsequently manipulate and clean the data.

```{r}
library(tidyverse)
library(RSocrata)
library(lubridate)
```

```{r, eval = FALSE}
## Install the required package with:
## install.packages("RSocrata) if not already installed

nyc_crime = read.socrata(
  "https://data.cityofnewyork.us/resource/9s4h-37hy.json",
  app_token = "jK8jo5IhF6gBPFqZ1va8ITny0",
  email     = "yakubuamin@ymail.com",
  password  = "daTA@12345" )

```

The acquired data has 6,036,805 observations and 35 variables. Broadly, the variables contain information on the exact date, time and location of crime, description of crime, demographic information of the victim and suspect and police infromation. For more information on the variables, click [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i).

### Subsetting 

We are interested in 2017, 2016, 2015 and 2014 so we extract crime information for those years. We also interested in felony crimes so take out misdeamenors and violations. 

```{r, eval = FALSE}
# This code was used to retrieve and assess the observations with missing date of crime occurrence

crime_missing_cp = nyc_crime %>% 
  summarize_all(funs(sum(is.na(.))))

#Cleaning and subsetting the data

nyc_felony_crimes = nyc_crime %>% 
  janitor::clean_names() %>% 
  mutate(year = year(cmplnt_fr_dt))%>%
  mutate_if(is.character, tolower) %>% 
  filter(year %in% 2014:2017) %>% 
  filter(law_cat_cd == "felony") %>% 
  select(- station_name, - transit_district, - hadevelopt, - patrol_boro, - housing_psa, - juris_desc)

#### selecting crimes of interest
sex_drug_weapons = crime_df %>% 
  filter(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 164, 179, 155, 586, 696, # Sex related felony crimes
                      ## Drug related felony crimes
                      500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 520, 521, 523, 524, 529, 530, 531, 532, 568, 570,
                      ### Weapon related felony crimes
                      781, 792, 793, 796)) %>% 
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>% 
  select(boro_nm, cmplnt_fr_dt, cmplnt_fr_tm, latitude, longitude, ky_cd, ofns_desc, pd_cd, pd_desc, vic_race, vic_sex, year, prem_typ_desc) %>% 
  mutate(boro_nm = if_else(boro_nm == "staten island", "staten_island", boro_nm),
         crime_group = if_else(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 164, 179, 155, 586, 696), "Sex-Related", 
                               if_else(pd_cd %in% c(500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 520, 521, 523, 524, 529, 530, 531, 532, 568, 570), "Drug-Related", 
                                       if_else(pd_cd %in% c(781, 792, 793, 796), "Weapon-Related", pd_cd))))

saveRDS(nyc_felony_crimes, file = "./datasets/nyc_felony_crimes.rds")
saveRDS(sex_drug_weapons, file = "./datasets/sex_drug_weapons.rds")
```

Our dataset for analysis has 588,704 observations and 31 variables for nyc_felony_crimes and it's subset sex_drug_weapons has 46,692 observations and 14 variables. 

To use our dataset, use the following code after cloning our repository. 

```{r, eval = FALSE}

### nyc_felony_crimes dataset
readRDS(file = "./datasets/nyc_felony_crimes.rds")

### sex_drug_weapons dataset
readRDS(file = "./datasets/sex_drug_weapons.rds")

sex_drug_weapons %>% 
  summarize_all(funs(sum(is.na(.))))
```

