---
title: "NYC Crime Data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

### Accessing the data. 

The data was accessed through the SODA Consumer API on 17th November, 2018. Click [here](https://dev.socrata.com/consumers/getting-started.html) to read more about the API. 

The `tidyverse`, `Rsocrata` and `lubridate` libraries are required to connect with the API and subsequently manipulate and clean the data.

```{r}
library(tidyverse)
library(RSocrata)
library(lubridate)
library(photon)
```

```{r, eval = FALSE}
## Install the required package with:
## install.packages("RSocrata) if not already installed

nyc_crime = read.socrata(
  "https://data.cityofnewyork.us/resource/9s4h-37hy.json",
  app_token = "jK8jo5IhF6gBPFqZ1va8ITny0",
  email     = "yakubuamin@ymail.com",
  password  = "daTA@12345" )

```

The acquired data has 6,036,805 observations and 35 variables. Broadly, the variables contain information on the exact date, time and location of crime, description of crime, demographic information of the victim and suspect and police infromation. For more information on the variables, click [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i).

### Subsetting 

We are interested in 2017, 2016, 2015 and 2014 so we extract crime information for those years. We also interested in felony crimes so take out misdeamenors and violations. 

```{r, eval = FALSE}
# This code was used to retrieve and assess the observations with missing date of crime occurrence

crime_missing_cp = nyc_crime %>% 
  summarize_all(funs(sum(is.na(.))))

#Cleaning and subsetting the data

nyc_felony_crimes = nyc_crime %>% 
  janitor::clean_names() %>% 
  mutate(year = year(cmplnt_fr_dt))%>%
  mutate_if(is.character, tolower) %>% 
  filter(year %in% 2014:2017) %>% 
  filter(law_cat_cd == "felony") %>% 
  select(- station_name, - transit_district, - hadevelopt, - patrol_boro, - housing_psa, - juris_desc)

#### selecting crimes of interest
sex_drug_weapons = nyc_felony_crimes %>% 
  filter(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 164, 179, 155, 586, 696, # Sex related felony crimes
                      ## Drug related felony crimes
                      500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 520, 521, 523, 524, 529, 530, 531, 532, 568, 570,
                      ### Weapon related felony crimes
                      781, 792, 793, 796)) %>% 
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>% 
  select(cmplnt_num, boro_nm, cmplnt_fr_dt, cmplnt_fr_tm, latitude, longitude, ky_cd, ofns_desc, pd_cd, pd_desc, vic_race, vic_sex, year, prem_typ_desc) %>% 
  mutate(boro_nm = if_else(boro_nm == "staten island", "staten_island", boro_nm),
         crime_group = if_else(pd_cd %in% c(178, 694, 697, 176, 180, 153, 157, 177, 168, 159, 166, 164, 179, 155, 586, 696), "Sex-Related", 
                               if_else(pd_cd %in% c(500, 501, 502, 503, 505, 507, 510, 512, 514, 515, 519, 520, 521, 523, 524, 529, 530, 531, 532, 568, 570), "Drug-Related", 
                                       if_else(pd_cd %in% c(781, 792, 793, 796), "Weapon-Related", pd_cd))))

saveRDS(nyc_felony_crimes, file = "./datasets/nyc_felony_crimes.rds")
saveRDS(sex_drug_weapons, file = "./datasets/sex_drug_weapons.rds")
```

Our dataset for analysis has 588,704 observations and 31 variables for nyc_felony_crimes and it's subset sex_drug_weapons has 46,692 observations and 14 variables. 

### Reverse Geocoding

The dataset comes without zipcode or neighborhood information. There are only geographic coordinates. This makes chloropleth mapping very difficut. We will reverse geocode the coordinates. 

### Felony Drug offenses

```{r}
#### Sampling the just year 2017 and selecting all felony drug rape reports

### Drug offenses in 2017
drug_2017 = sex_drug_weapons %>% filter(crime_group == "Drug-Related" & year == 2017)

### Sampling to reduce the dataset so has not not get rejected by google API
drug_2017_1 = drug_2017 %>% sample_n(2290)

drug_2017_zip_1 = drug_2017_1 %>% 
  mutate(zip = map2(.x = longitude, .y = latitude, ~ revgeocode(c(lon = .x, lat = .y))))

saveRDS(drug_2017_zip_1, file = "datasets/drug_2017_zip_1.rds")

#### Getting the data not sample. Second half of the data
drug_2017_2 = anti_join(drug_2017, drug_2017_zip_1, by = "cmplnt_num")

#### Iterating to get the data
drug_2017_zip_2 = drug_2017_2 %>% 
  mutate(zip = map2(.x = longitude, .y = latitude, ~ revgeocode(c(lon = .x, lat = .y))))

saveRDS(drug_2017_zip_2, file = "datasets/drug_2017_zip_2.rds")

### Joining to have the complete dataset

complete_drug_zip_2017 = bind_rows(drug_2017_zip_1, drug_2017_zip_2) 

saveRDS(complete_drug_zip_2017, file = "datasets/complete_drug_zip_2017.rds")
```

### Sex Drug offenses

### Usage
To use our dataset, use the following code after cloning our repository. 

```{r, eval = FALSE}

### nyc_felony_crimes dataset
readRDS(file = "./datasets/nyc_felony_crimes.rds")

### sex_drug_weapons dataset
readRDS(file = "./datasets/sex_drug_weapons.rds")

### Drug offenses for 2017 with zipcodes
saveRDS(complete_drug_zip_2017, file = "datasets/complete_drug_zip_2017.rds")

```

