---
title: "Exploratory Analysis"
author: "Jyoti Ankam"
date: "November 17, 2018"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

Loading the relevant libraries

```{r}
library(tidyverse)
library(rvest)
library(plotly)
library(sqldf)
```

Reading the dataset

```{r}

crime_df = readRDS(file = "datasets/nyc_felony_crimes.rds")

```

Loading information from the US census on total population in the individual NYC boroughs to calculate offense rates in each borough
```{r}

url = "https://www.census.gov/quickfacts/fact/table/newyorkcitynewyork,bronxcountybronxboroughnewyork,kingscountybrooklynboroughnewyork,newyorkcountymanhattanboroughnewyork,queenscountyqueensboroughnewyork,richmondcountystatenislandboroughnewyork/PST045217"
boro_pop = read_html(url)

boro_pop %>%
  html_nodes(css = "table")

pop_df = (boro_pop %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table() %>% 
  as_tibble() %>% 
 janitor::clean_names()

names(pop_df)[1:7] = c("estimate_date", "new_york_city", "bronx", "brooklyn", "manhattan", "queens", "staten_island")

pop_df = pop_df %>% 
  gather(key = boro_nm, value = population, estimate_date:staten_island) %>% 
  mutate(population = if_else(population == "Population estimates, July 1, 2017,  (V2017)", "2017", population),
         population = as.numeric(gsub("," , "", population)))

```

```{r}
nycc_df = crime_df %>% 
  mutate(boro_nm = if_else(boro_nm == "staten island", "staten_island", boro_nm))
```
 
Looking at the individual offenses by borough over the four years:
```{r}

crime_df %>% 
  distinct(pd_cd) %>% 
  count()

combined_df = nycc_df %>% group_by(boro_nm, pd_cd, year) %>% 
  summarise(count = n())


full = left_join(combined_df, pop_df, by = "boro_nm") %>% 
mutate(off_rate = ((count/population)*100)) 

full[,'off_rate'] = round(full[,'off_rate'],2)

test <- sqldf("SELECT boro_nm, year, off_rate
              FROM full
              WHERE off_rate IS NOT NULL and year = '2017'
              ORDER BY boro_nm, off_rate DESC")

test_2017 <- test[!duplicated(test$boro_nm), ]


test2 <- sqldf("SELECT boro_nm, year, off_rate
              FROM full
              WHERE off_rate IS NOT NULL and year = '2016'
              ORDER BY boro_nm, off_rate DESC")

test_2016 <- test2[!duplicated(test2$boro_nm), ]


test3 <- sqldf("SELECT boro_nm, year, off_rate
              FROM full
              WHERE off_rate IS NOT NULL and year = '2015'
              ORDER BY boro_nm, off_rate DESC")

test_2015 <- test3[!duplicated(test3$boro_nm), ]


test4 <- sqldf("SELECT boro_nm, year, off_rate
              FROM full
              WHERE off_rate IS NOT NULL and year = '2014'
              ORDER BY boro_nm, off_rate DESC")

test_2014 <- test4[!duplicated(test4$boro_nm), ]

test_new = rbind(test_2017, test_2016, test_2015, test_2014)
 

trend = ggplot(test_new, aes(x = boro_nm, y = off_rate, color = year)) +
geom_point() + 
theme(legend.position = "bottom") +
  labs(
  x = "Borough",
  y = "Offense rate",
  caption = "NYC_Crime"
  ) +
geom_smooth(se = FALSE) +
theme(legend.position = "bottom")

ggplotly(trend)

```

Interpretation: From this exploratory subanalysis, we can see that the rates of individual offenses are highest in the boroughs of Bronx and Brooklyn, followed by Manhattan and Queens. Staten Island reported the lowest rates of individual offenses throughout the four years. Individual offense rates increased in the Bronx from 2014 through 2017. On the contrary, Brooklyn witnessed a significant plummeting in offense rates through the years 2014 to 2017.