---
title: "date_and_borough"
author: "Jacky Choi"
date: "November 18, 2018"
output: github_document
---

\
This is one of the exploratory analyses for our P8105 - Data Science group project.

\
First, I load the tidyverse and patchwork package.
```{r setup, echo = FALSE}
library(tidyverse)
devtools::install_github("thomasp85/patchwork")
library(patchwork)
```

\
Next, I read in the CSV file and tidy it up. I limit the data to just individuals in the borough of Queens and randomly sample 500 observations from the data frame.
```{r tidy_crime_data}
crime_df = readRDS(file = "datasets/nyc_felony_misdemeanor.rds")

crime_data = readRDS(file = "datasets/nyc_felony_misdemeanor.rds") %>% 
  janitor::clean_names() %>%
  sample(nrow(crime_df), 500) %>% 
  filter(boro_nm == "Queens") %>% 
  separate(cmplnt_fr_dt, into = c("year_of_complaint", "month_complaint", "day_complaint"), sep = "-") %>% 
  select(cmplnt_fr_dt, law_cat_cd, boro_nm)

head(crime_data)
```

\
I can look at the data in a table to examine general trends and get a feel of my data.
```{r crime_data_table}
crime_table = 
  crime_data %>% 
  knitr::kable(digits = 2)

head(crime_table, 12)
```

\
It is hard to visualize any trends using solely the data table, so I now attempt to graph my data. I want to graph the year of the complaint against the category of offense in Queens. That is, I want to see if any particular year in Queens was associated with a higher category of offense.
```{r graph_crime_data}
crime_graph = 
  crime_data %>% 
  ggplot(aes(x = year_of_complaint, y = law_cat_cd)) + 
  geom_point() +
    labs(
    title = "Categories of Offenses, By Year of Complaint, In Queens",
    x = "Year",
    y = "Category",
    caption = "Data from the crime_data file"
    ) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

crime_graph
```

I can also look at whether the length of time for the reported event/crime differs by borough.
