---
title: "Exploratory Analyses"
output:
  html_document:
    toc: true
    toc_float: true
---

<br>

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: red;
}

.navbar-default .navbar-collapse, .navbar-default .navbar-form {
background-color: red;

</style>


## Exploratory Analysis 1 - Length of Reported Event/Crime by Borough

<br>
As one of our exploratory analyses, we thought it would be interesting to examine whether the length of time of the reported event or crime differs by borough. Examining the length of the reported event or crime can serve as a proxy indicator of the severity of the crime by each borough, assuming that longer crimes tend to be more severe, harder to resolve, and requiring more resources to deal with.
```{r exp_setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r exp_libraries}
library(patchwork)
library(tidyverse)
library(rvest)
library(plotly)
library(sqldf)
library(tidyverse)
library(htmltools)
library(htmlwidgets)
library(leaflet)
library(leaflet.extras)
library(chron)
library(lubridate)
library(rgdal)
if (!require(geojsonio)) {
    install.packages("geojsonio")
    library(geojsonio)
}
library(sp)
library(maps)
library(ggmap)
library(maptools)
library(sp)
install.packages("tmap")
library(tmap)
```

<br>
First, we loaded the required R packages. Next, we read in the CSV file and tidied it up. For this exploratory analysis, we chose to limit the data by randomly sampling 1,000 observations from the data frame to simply the analysis a bit. We also chose to include only observations with non-missing values to the following studied variables: cmplnt_to_dt (start date of reported crime/event), cmplnt_fr_dt (end date of reported crime/event), and law_cat_cd (classification of crime/event; i.e., felony). The number of observations was large and sufficient enough to observe any patterns - our final analytic sample included 861 observations.
```{r exp_tidy_crime_data, echo = FALSE}
crime_df = readRDS(file = "./datasets/nyc_felony_crimes.rds")

crime_data = crime_df %>% 
  janitor::clean_names() %>%
  sample_n(1000) %>% 
  mutate(time_diff = (difftime(cmplnt_to_dt, cmplnt_fr_dt))) %>% 
  mutate(time_diff2 = (as.numeric(cmplnt_to_dt - cmplnt_fr_dt, units = "days", na.rm = TRUE))) %>% 
  select(cmplnt_fr_dt, cmplnt_to_dt, law_cat_cd, time_diff, time_diff2, boro_nm)

crime_data2 <- crime_data[complete.cases(crime_data),]

```

<br>
We looked at the data in a data table to examine general trends and get a feel of the data. Our data cleaning seemed to work.
```{r exp_crime_data_table, echo = FALSE}
crime_table = 
  crime_data2 %>% 
  knitr::kable(digits = 2)

head(crime_table)
```

<br>
Next, we looked at whether the length of time for the reported event/crime differed by borough visually.
```{r exp_length_of_time_of_reported_event, echo = FALSE}
crime_data2 <- crime_data[complete.cases(crime_data),]

crime_graph2 = 
  crime_data2 %>% 
  group_by(boro_nm, law_cat_cd) %>% 
  ggplot(aes(x = boro_nm, y = time_diff2, color = law_cat_cd)) + 
  geom_point() +
    labs(
    title = "Spread of Times of Reported Event, By Borough",
    x = "Borough",
    y = "Length of Reported Event",
    caption = "Data from the crime_data file"
    ) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

crime_graph2
```

<br>
From the graph above, we can see a couple of things. First, there are, numerically speaking, fewer reported events in Staten Island. Secondly, the spread of the reported events seems to be slightly larger in Queens than in other boroughs. Lastly, although the spread among Brooklyn and Manhattan is approximately the same, Queens has more outliers.

<br>
You can find the code for this exploratory analysis [here](code_date_borough.html).

<br>
<br>

## Exploratory Analysis 2 - Overall Crime Rates by Borough

For our second exploratory analysis, we chose to examine crime rates by borough. Doing so gives us an understanding as to how NYPD can prioritize their efforts when stemming crime.

First, we load the required R packages and read in the dataset. We also load information from the U.S. Census on total population in the individual NYC boroughs to calculate offense rates in each borough.

```{r}
grouped_df = sex_drug_weapons %>% group_by(boro_nm, year) %>% 
  summarise(number = n())

full = left_join(grouped_df, nyc_population, by = "boro_nm") %>% 
  mutate(crime_rate = (number/population)*100)

x = full %>% filter(!is.na(boro_nm)) %>% 
  ggplot(aes(x = year, y = crime_rate, color = boro_nm)) + 
  geom_point() + geom_line(size = 1) +
  labs(x = "Year",
       y = "Crime rate",
       legend) + viridis::scale_color_viridis(
      name = "Borough", 
      discrete = TRUE
    ) + theme_classic()

ggplotly(x)
```


<br>
<br>

## Exploratory Analysis 3 - Top Crimes and Associated Demographics by Borough
Include Yaa's

<br>
<br>

## Exploratory Analysis 4 - Geographical Distribution of Crimes
Reading in the datasets

```{r, exploratory_data}
###Dataset for 2017 with addresses
crimes_df = readRDS(file = "./datasets/all_sex_drug_weapon_zip_2017.rds")

### Dataset for 4 years without addresses
crimes_14_17 = readRDS(file = "./datasets/sex_drug_weapons.rds")

### Selecting zipcodes from address
crimes_df = crimes_df %>% 
  mutate(zipcode = str_extract(zip, "[0-9][0-9][0-9][0-9][0-9]"))

```

```{r, exp_leaflet}
year_2014 = filter(crimes_14_17, year == 2014) %>% 
  filter(!is.na(longitude) | !is.na(latitude))

year_2015 = filter(crimes_14_17, year == 2015) %>% 
  filter(!is.na(longitude) | !is.na(latitude))

year_2016 = filter(crimes_14_17, year == 2016) %>% 
  filter(!is.na(longitude) | !is.na(latitude))

year_2017 = filter(crimes_14_17, year == 2017) %>% 
  filter(!is.na(longitude) | !is.na(latitude))

pal <- colorFactor(palette = c("blue", "red", "green"), 
               levels = c("Sex-Related", "Weapon-Related", "Drug-Related"))


leaflet()  %>% 
  addTiles() %>% 
#Setting view on NYC. The values are the coordinates of NYC
    setView(lat = 40.7, lng = -74.0, zoom = 11) %>%
  addCircleMarkers(data = year_2014,
                   lng = ~ longitude,
                   lat = ~ latitude, 
                   radius = 1,
                   color = ~pal(crime_group),
                   group = "2014",
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = year_2015,
                   lng = ~ longitude,
                   lat = ~ latitude, 
                   radius = 1,
                   color = ~pal(crime_group),
                   group = "2015",
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = year_2016,
                   lng = ~ longitude,
                   lat = ~ latitude, 
                   radius = 1,
                   color = ~pal(crime_group),
                   group = "2016",
                   clusterOptions = markerClusterOptions()) %>% 
  addCircleMarkers(data = year_2017,
                   lng = ~ longitude,
                   lat = ~ latitude, 
                   radius = 1,
                   color = ~pal(crime_group),
                   group = "2017",
                   clusterOptions = markerClusterOptions()) %>%
  addLayersControl(overlayGroups = c("2014", "2015", "2016", "2017")) %>% 
   addLegend(pal = pal, 
            values = c("Sex-Related", "Weapon-Related", "Drug-Related"),
            opacity = .5,
            title = "Crime Group",
            position = "topleft")
```

```{r, eval = FALSE}

map_url = "http://data.beta.nyc//dataset/3bf5fb73-edb5-4b05-bb29-7c95f4a727fc/resource/6df127b1-6d04-4bb7-b983-07402a2c3f90/download/f4129d9aa6dd4281bc98d0f701629b76nyczipcodetabulationareas.geojson"

download.file(map_url, "./datasets/nyc_zip_map.geojson")
```

#### Crime distribution by zipcode for 2017

```{r, exp_zips}

group_zip = crimes_df %>% group_by(zipcode) %>% summarize(number = n())

map_json <- geojson_read("./datasets/nyc_zip_map.geojson", what = "sp")

map_w_data = sp::merge(map_json, group_zip, by.x = "postalCode", by.y = "zipcode") 

map_w_data@data <-  
  map_w_data@data %>% 
  mutate(number = as.numeric(number)) %>% 
  mutate(number = if_else(is.na(number), 0, number))

total = tm_shape(map_w_data) +
  # Add title and change palette
  tm_fill(col = "number", 
          title = "Total Crime",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5) 

```

```{r, exp_drug}
drug_group = crimes_df %>% filter(crime_group == "Drug-Related") %>% 
  group_by(zipcode) %>% 
  summarize(number = n())

map_json_drug <- geojson_read("./datasets/nyc_zip_map.geojson", what = "sp")

map_w_drug = sp::merge(map_json_drug, drug_group, by.x = "postalCode", by.y = "zipcode") 

map_w_drug@data <-  
  map_w_drug@data %>% 
  mutate(number = as.numeric(number)) %>% 
  mutate(number = if_else(is.na(number), 0, number))

# Use qtm() to create a choropleth map of gdp
drug = tm_shape(map_w_drug) +
  # Add title and change palette
  tm_fill(col = "number", 
          title = "2017 Drug-Related Crimes",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5)

```


```{r, exp_weapon}
weapon_group = crimes_df %>% filter(crime_group == "Weapon-Related") %>% 
  group_by(zipcode) %>% 
  summarize(number = n())

map_json_weapon <- geojson_read("./datasets/nyc_zip_map.geojson", what = "sp")

map_w_weapon = sp::merge(map_json_drug, weapon_group, by.x = "postalCode", by.y = "zipcode") 

map_w_weapon@data <-  
  map_w_weapon@data %>% 
  mutate(number = as.numeric(number)) %>% 
  mutate(number = if_else(is.na(number), 0, number))

# Use qtm() to create a choropleth map of gdp

weapon = tm_shape(map_w_weapon) +
  # Add title and change palette
  tm_fill(col = "number",
          title = "2017 Weapon-Related Crimes",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5)
  # Add tm_credits()
```


```{r, exp_sex}
sex_group = crimes_df %>% filter(crime_group == "Sex-Related") %>% 
  group_by(zipcode) %>% 
  summarize(number = n())

map_json_sex <- geojson_read("./datasets/nyc_zip_map.geojson", what = "sp")

map_w_sex = sp::merge(map_json_sex, sex_group, by.x = "postalCode", by.y = "zipcode") 

map_w_sex@data <-  
  map_w_sex@data %>% 
  mutate(number = as.numeric(number)) %>% 
  mutate(number = if_else(is.na(number), 0, number))

# Use qtm() to create a choropleth map of gdp
sex = tm_shape(map_w_sex) +
  # Add title and change palette
  tm_fill(col = "number",
          title = "2017 Sex-Related Crimes",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5) 

current.mode <- tmap_mode("plot")
tmap_arrange(total, sex, drug, weapon)
tmap_mode(current.mode)
```

#### Trend

```{r exp_analysis_pop}
url = "https://www.census.gov/quickfacts/fact/table/newyorkcitynewyork,bronxcountybronxboroughnewyork,kingscountybrooklynboroughnewyork,newyorkcountymanhattanboroughnewyork,queenscountyqueensboroughnewyork,richmondcountystatenislandboroughnewyork/PST045217"

nyc_population = read_html(url) %>%  html_nodes(css = "table") %>% .[[1]] %>% 
  html_table(header = TRUE) %>% as.tibble() %>% 
  janitor::clean_names()

names(nyc_population)[1:7] = c("estimate_date", "new_york_city", "bronx", "brooklyn", "manhattan", "queens", "staten_island")

nyc_population = nyc_population %>% 
  gather(key = boro_nm, value = population, estimate_date:staten_island) %>% 
  mutate(population = if_else(population == "Population estimates, July 1, 2017,  (V2017)", "2017", population),
         population = as.numeric(gsub("," , "", population)))

```

```{r exp_trend_analysis_stratify}
stratify_df = crimes_14_17 %>% group_by(boro_nm, crime_group, year) %>% 
  summarise(number = n())

population_df = left_join(stratify_df, nyc_population, by = "boro_nm") %>% 
  mutate(crime_rate = (number/population)*100)
```

```{r trend_analysis_graph}
population_df %>% filter(!is.na(boro_nm)) %>% 
  ggplot(aes(x = year, y = crime_rate, color = boro_nm)) + 
  geom_point() + geom_line(size = 1) + facet_grid(~crime_group) +
  labs(x = "Year",
       y = "Crime rate",
       legend) + viridis::scale_color_viridis(
      name = "Borough", 
      discrete = TRUE
    )
```

Bronx has consistently had the highest crime rate over the last 4 years for all the 3 types of crimes. We see a steep reduction, however in the crime rate for bronx between the years of 2016 and 2017. Crime rates for all crimes in all boroughs decreased from 2016 to 2016. Surprisingly, brooklyn has seen a consitent rise in the crime rate of drug related felonies from 2015 to 2017. 

#### Timing




