---
title: "Final_project"
author: "Yaa Klu"
date: "11/17/2018"
output: github_document
---

```{r}
crime_nyc = readRDS(file = "datasets/nyc_felony_misdemeanor.rds")
```
 

```{r}
library(tidyverse)
```


**Crime category by race/ethnicity, age, sex and other sociodemographics**
** top ten offense, general chracteristics**
** top 5, general characteristics**
**filter out unknown and analyze top 5 crimes for unknowns**


Removing missing variables
```{r}
com_crime_nyc = crime_nyc %>% 
  janitor::clean_names(dat = .) %>% 
  filter(!is.na(susp_sex)) %>%
  filter(!is.na(susp_age_group)) %>%
  filter(!is.na(vic_sex)) %>%
  filter(!is.na(vic_age_group)) %>%
  filter(!is.na(vic_race)) %>%
  filter(!is.na(susp_race))
```


Highest number of crimes (top 5)
```{r}
com_crime_nyc %>%
  distinct(pd_desc) %>%
  count()

com_crime_nyc %>%
  filter(law_cat_cd == "felony") %>% 
  group_by(pd_desc, pd_cd) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(10, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```


Suspect Age
```{r}
com_crime_nyc %>%
  group_by(susp_age_group) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```

Suspect Sex
```{r}
com_crime_nyc %>%
  group_by(susp_sex) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(3, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 

```

Victim's sex
```{r}
com_crime_nyc %>%
  group_by(vic_sex) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(3, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```

Victim's age
```{r}
com_crime_nyc %>%
  group_by(vic_age_group) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```

Suspect's age
```{r}
com_crime_nyc %>%
  group_by(susp_race) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```


Victim's race
```{r}
com_crime_nyc %>%
  group_by(vic_race) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable()
```


Suspect's race
```{r}
com_crime_nyc %>%
  group_by(pd_desc, susp_race) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable()

```


```{r}
com_crime_nyc %>%
  group_by(susp_race, vic_race) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable()
```


```{r}
com_crime_nyc %>%
  group_by(susp_race, pd_desc) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable()
```


```{r}
com_crime_nyc %>%
  group_by(pd_desc, susp_age_group) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(5, number) %>%
  arrange(desc(number)) %>%
  knitr::kable() 
```


**filter out unknown and analyze top 5 crimes for unknowns**
```{r}
filter(com_crime_nyc, susp_age_group == 'unknown') %>%
  group_by(pd_cd) %>%
  summarise(n = n()) %>%
  filter() %>%
  knitr::kable() 
```


select top 10 crimes into a new dataset
```{r}
top_ten_crime = com_crime_nyc %>%
  group_by(pd_desc, pd_cd) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  top_n(10, number) 
 # arrange(desc(number)) %>%
```

```{r}
mydataextract3 = list()
cc = top_ten_crime$pd_cd
for(j in 1:length(cc)){
  ind = cc[j]
#mydataextract =  com_crime_nyc[com_crime_nyc$pd_desc == cc, ]
mydataextract3[[j]] = com_crime_nyc[which(com_crime_nyc$pd_cd == ind, arr.ind = TRUE ), ]
mydataextract4 = do.call(rbind, mydataextract3)
} # has 397,023 obs
```


consider suspects demographics by the top ten crimes
suspect age group
```{r}
 table(mydataextract4$susp_age_group)
sort(table(mydataextract4$susp_age_group))
#keep age groups < 18, 18-24, 25-44, 45-64, 65+, unknown
agegroup=c('<18', '45-64',   '18-24', 'unknown',   '25-44', '65+' )
mydata=mydataextract4[mydataextract4$susp_age_group %in% agegroup,]
suspect_agetopten=table(mydata$pd_desc,mydata$susp_age_group)
```



```{r}
suspect_agetopten2=as.data.frame(suspect_agetopten)
mycolnames=c('Crime', 'Agegroup', 'Freq')
colnames(suspect_agetopten2)=mycolnames
suspect_agetopten2 %>%
group_by(Agegroup, Crime) %>%
  ggplot(aes(x = Crime, y = Freq, color = Agegroup)) +
   geom_point() +
#   geom_smooth(se=FALSE) +
labs(
  title = "Graph of crime frequency by age group",
  x = "Crime ",
  y = "Freq"
) +
  theme(axis.text.x = element_text(angle = 45, size = 7, hjust = 1))

```


consider suspects demographics by the top ten crimes.
suspect race
```{r}
 table(mydataextract4$susp_race)
sort(table(mydataextract4$susp_race))
#keep age groups < 18, 18-24, 25-44, 45-64, 65+, unknown
suspect_racetopten=table(mydata$pd_desc,mydata$susp_race)
```



```{r}
suspect_racetopten2=as.data.frame(suspect_racetopten)
mycolnames=c('Crime', 'Race', 'Freq')
colnames(suspect_racetopten2)=mycolnames
suspect_racetopten2 %>%
group_by(Race, Crime) %>%
  ggplot(aes(x = Crime, y = Freq, color = Race)) +
   geom_point() +
#   geom_smooth(se=FALSE) +
labs(
  title = "Graph of crime frequency by race",
  x = "Crime ",
  y = "Freq"
) +
  theme(axis.text.x = element_text(angle = 45, size = 7, hjust = 1))

```


sex
```{r}
table(mydataextract4$susp_sex)
sort(table(mydataextract4$susp_sex))
#keep age groups < 18, 18-24, 25-44, 45-64, 65+, unknown
suspect_sextopten=table(mydata$pd_desc,mydata$susp_sex)
suspect_sextopten
```

```{r}
suspect_sextopten2=as.data.frame(suspect_sextopten)
mycolnames=c('Crime', 'Sex', 'Freq')
colnames(suspect_sextopten2)=mycolnames
suspect_sextopten2 %>%
group_by(Sex, Crime) %>%
  ggplot(aes(x = Crime, y = Freq, color = Sex)) +
   geom_point() +
#   geom_smooth(se=FALSE) +
labs(
  title = "Graph of crime frequency by sex",
  x = "Crime ",
  y = "Freq"
) +
  theme(axis.text.x = element_text(angle = 45, size = 7, hjust = 1))

```

